A Large-Scale Study of the Use of Eval in R

### 2020-08-24 inst/analysis/corpus-analysis.Rmd
V.A

  r("code",                 sum(corpus$package_code)),
  r("R code",               sum(corpus$package_r_code)),
  r("native code",          sum(corpus$package_native_code)),

@@

```{r generals}
overview_table(
  r("all cran", dim(available.packages())[1]),
  r("core packages", length(CORE_PACKAGES))
)
```

@@

corpus %>%
  mutate(
    outlier=package_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(rank <= 25 | outlier, package, NA)  #if_else(outlier, str_glue("{package}"), as.character(NA))
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
      y=revdeps,
      label=label,
      size=package_code
    )
  ) +
  geom_point(alpha=.6) +
  geom_text(size=2, check_overlap = T, vjust=1.5, na.rm = TRUE) +

@@

```{r}
kaggle_runnable <- filter(kaggle, runnable)
kaggle_finished <- filter(kaggle, exitval==0)
kaggle_code     <- kaggle_finished$code
kaggle_runtime  <- kaggle_finished$runtime

@@

```{r}
overview_table(
  r("all programs", nrow(kaggle_finished) + sum(corpus_programs_files)),
  r("all runnbale code", sum(kaggle_code) + sum(runnable_code))
)
```

```{r}
corpus_programs <-
  left_join(
    filter(runs, exitval==0) %>% select(package, file, time), 
    runnable_code_metadata %>% select(package, file, code),
    by=c("package", "file")
  ) %>%
  transmute(id=paste(package, ":", file), code, runtime=time, corpus="packages")

>>>>

  r("code",                 sum(corpus$package_code)),
  r("R code",               sum(corpus$package_r_code)),
  r("R code mean",          mean(corpus$package_r_code)),
  r("R code median",        median(corpus$package_r_code)),
  r("native code",          sum(corpus$package_native_code)),
  r("native code mean",     mean(corpus$package_native_code)),
  r("native code median",   median(corpus$package_native_code)),

@@

```{r generals}
overview_table(
  r("all cran", dim(available.packages())[1]),
  r("core packages", length(CORE_PACKAGES)),
  r("most downloaded treshold", MOST_DOWNLOADED),
  r("large package treshold", PACKAGE_SIZE_OUTLIER)
)
```

@@

corpus %>%
  mutate(
    outlier=package_code > PACKAGE_SIZE_OUTLIER,
    label=ifelse(rank <= MOST_DOWNLOADED | outlier, package, NA)  #if_else(outlier, str_glue("{package}"), as.character(NA))
  ) %>%
  ggplot(
    aes(
      x=coverage_expr, 
      y=revdeps,
      label=label,
      size=package_code
    )
  ) +
  geom_point(alpha=.6) +
  geom_text(size=2, check_overlap = T, vjust=1.5, na.rm = TRUE) +

@@

```{r kaggle summary}
kaggle_runnable <- filter(kaggle, runnable)
kaggle_finished <- filter(kaggle, exitval==0)
kaggle_code     <- kaggle_finished$code
kaggle_runtime  <- kaggle_finished$runtime

@@

```{r all summary}
overview_table(
  r("all programs", nrow(kaggle_finished) + sum(corpus_programs_files)),
  r("all runnbale code", sum(kaggle_code) + sum(runnable_code)),
  r("all code", sum(kaggle_code) + sum(corpus$package_code) + sum(runnable_code))
)
```

```{r compute programs}
corpus_programs <-
  left_join(
    filter(runs, exitval==0) %>% select(package, file, time), 
    runnable_code_metadata %>% select(package, file, code),
    by=c("package", "file")
  ) %>%
  transmute(id=paste(package, ":", file), code, runtime=time, corpus="packages")

### 2020-08-24 inst/paper/main.tex
V.A

We have assembled a corpus that consists of \CorpusPackages R packages and user programs.

We have selected \CorpusPackages packages consisting of \CorpusRCodeRnd
lines of R code and \CorpusNativeCodeRnd lines of native code
(C/C++/Fortran).  Figure~\ref{fig:corpus} shows these packages, the size of
the dots reflects the project's size in lines of code including both R and
native code\footnote{Excluding comments and blank lines,
  \url{github.com/AlDanial/cloc}}, the x-axis indicates the expression code
coverage in percents and the y-axis gives the number of reverse dependencies
in log scale. Dotted lines indicate means. Large and popular packages are
annotated.

@@

These packages come from the Comprehensive R Archive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \CorpusAllCranRnd packages\footnote{CRAN receives about 6 new
  package submissions a day~\cite{Ligges2017}}. Unlike other open code
repositories such as GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are automatically
checked asserting certain quality. Most relevant for this work is that all of
the runnable code (including code snippets from examples and vignettes) is
tested and only a successfully running package is admitted in the archive.

\begin{figure}[!tb]\centering\includegraphics[width=.9\linewidth]
  {programs-code.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure}
\begin{figure}[!tb]\centering\includegraphics[width=.9\linewidth]
  {programs-runtime.pdf}\caption{Corpus}\label{fig:corpus}
\end{figure}

% FILIP: WIP
% We have downloaded and installed all available CRAN packages. Out of the
% \AllCranRnd packages, we managed to install \AllLoadableRnd. The main reason for
% this is that \rdt is based on GNU-R 3.5.0 and some of the packages are no longer
% compatible with this version. Some packages also require extra native
% dependencies which were not present on our servers. We defined two criteria for
% including a package into the corpus:
% \begin{inparaenum}[(1)]
%   \item have a runnable code that covers a significant part of the package
%   source code from which type signatures could be inferred, and
%   \item have some reverse dependencies that will allows us to evaluate the
%   inferred types using the runnable code from these dependencies.
% \end{inparaenum}
% The concrete thresholds used were: at least \ThresholdCodeCoverage of expression
% coverage and at minimum \ThresholdRevdeps reverse dependencies. The code
% coverage was computed for each package using
% \covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R tool.
% The reverse package dependencies were extracted from the package metadata using
% builtin function.

% The \CorpusLoadable selected packages contain \CorpusRunnableCodeRnd lines of
% runnable code in examples (\CorpusExamplesCodeRnd), tests (\CorpusTestsCodeRnd)
% and vignettes (\CorpusVignettesCodeRnd). Running this code results on average in
% \CorpusMeanExprCoverage package code coverage (CRAN average is
% \AllMeanExprCoverage). Together, there is \CorpusRevdesRnd (on average
% \CorpusMeanRevdes, median \CorpusMedianRevdes; CRAN average is \AllMeanRevdes,
% median \AllMedianRevdes) reverse dependencies with \CorpusRevdepRunnableCodeRnd
% runnable lines of code resulting in \CorpusRevdepMeanCodeCoverage coverage (on
% average) of the corpus packages.
% Together there are \CorpusFunctionsRnd defined R functions
% (\CorpusPublicFunctionsRnd are from packages' public API).
% \CorpusSThreeFunctionsRnd are S3 functions. Packages in the corpus define
% \CorpusSThreeClasses S3 classes.
% The code coverage was computed for each package using
% \covr\footnote{\url{https://github.com/r-lib/covr}}, the R code coverage R
% tool.  The reverse package dependencies were extracted from the package
% metadata using builtin function.

>>>>

We have assembled a corpus of \CorpusPackages R packages and
\CorpusFinishedKaggle user-written data-analysis scripts. Together it contains
\CorpusAllCodeRnd lines of source code\footnote{Excluding comments and blank lines using \url{github.com/AlDanial/cloc}}.

\subsection{CRAN Packages}

We have selected the TOP \CorpusPackages packages based on the number of reverse
dependencies, \ie how many packages depend on them\footnote{extracted from
  package metadata using builtin function}. Using the number of reverse
dependencies helps us to choose popular packages which in general should be of a
better quality and should have better code coverage. The resulting set consists
of \CorpusRCodeRnd (median=\CorpusRCodeMedian) lines of R code and
\CorpusNativeCodeRnd (median=\CorpusNativeCodeMedian) lines of native code
(C/C++/Fortran). Figure~\ref{fig:corpus} shows these packages, the size of the
dots reflects the project's size in lines of code including both R and native
code, the x-axis indicates the expression code coverage in percents and the
y-axis gives the number of reverse dependencies in log scale. Dotted lines
indicate means. Large (over \CorpusLargePackageTresholdRnd lines of code) and popular (in \CorpusMostDownloadedTreshold most downloaded) packages are annotated.

@@

These packages come from the Comprehensive R Archive Network
(CRAN\footnote{\url{http://cran.r-project.org}}), the largest repository of R
code with over \CorpusAllCranRnd packages\footnote{CRAN receives about 6 new
  package submissions a day~\cite{Ligges2017}}. Unlike other open code
repositories such as GitHub, CRAN is a curated repository. Each submitted
package must abide to a number of well-formedness rules that are automatically
checked asserting certain quality.

Next to code and tests, R packages also contain examples and vignettes. Examples
are runnable snippets extracted from documentation. Vignettes are tutorial-like
guides into the package functionality. They are written in the literate
programming style using either Latex or Rmarkdown format with embedded code
snippets. Since CRAN is a curated repository, upon submission, it checks all the
runnable code in tests, examples and vignettes and only a successfully running
package is admitted to the archive. We use the same mechanism to extract
runnable code and run it for the dynamic analysis.

In the selected packages, there are \CorpusRunnableCodeRnd lines of runnable
code in examples (\CorpusExamplesCodeRnd), tests (\CorpusTestsCodeRnd) and
vignettes (\CorpusVignettesCodeRnd). Running this code results on average in
\CorpusMeanExprCoverage package code coverage. The code coverage was computed
for each package using

\begin{figure}[!tb]\centering\includegraphics[width=.9\linewidth]
  {programs-code.pdf}\caption{Programs code size}\label{fig:programs-code}
\end{figure}
\begin{figure}[!tb]\centering\includegraphics[width=.9\linewidth]
  {programs-runtime.pdf}\caption{Programs execution time}\label{fig:programs-runtime}
\end{figure}

\subsection{Kaggle sripts}
% contrast to CRAN - duplication
